---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r,  include = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
```


```{r, include = FALSE}
source("R/functions/optimise_gross_flows.R")
source("R/functions/split_gross_flows.R")
```

```{r, include = FALSE}
sel_code <- "E09000007"
sel_sex <- "female"
sel_components <- c("international_in", "international_out", "international_net", "unattrib")

mye_comp <- bind_rows(
  readRDS("data/intermediate/mye_2011_on(2023_geog).rds") %>%
    mutate(source = "original"),
  readRDS("data/processed/new_mye_series_2011_on.rds") %>%
    mutate(source = "adjusted")) %>%
  filter(gss_code == sel_code,
         sex == sel_sex,
         component %in% sel_components)

mye_comp_totals <- mye_comp %>%
  group_by(across(-any_of(c("value", "age")))) %>%
  summarise(value = sum(value), .groups = "drop") %>%
  arrange(year) %>%
  group_by(across(-any_of(c("value", "year")))) %>%
  mutate(cumulative_value = cumsum(value)) %>%
  ungroup()

```

```{r, include = FALSE}

sel_codes <- c("E08000035", "E06000018", "E08000026", "E09000007")
sel_sex <- "female"
sel_components <- c("international_in", "international_out", "international_net",
                    "unattrib", "total_net", 
                    "internal_in", "internal_out", "internal_net")

dpm_comp <- bind_rows(
  readRDS("data/intermediate/mye_2011_on(2023_geog).rds") %>%
    mutate(source = "MYE original"),
  readRDS("data/processed/new_mye_series_2011_on.rds") %>%
    mutate(source = "MYE adjusted"),
  readRDS("data/processed/new_dpm_series_2011_on.rds") %>%
    mutate(source = "DPM")) %>%
  filter(gss_code %in% sel_codes,
         sex == sel_sex,
         component %in% sel_components) 


dpm_total_flow_comps <- dpm_comp %>%
  filter(component %in% c("international_in", "international_out", "international_net",
                           "internal_in", "internal_out", "internal_net")) %>%
  group_by(across(-any_of(c("value", "component")))) %>%
  summarise(total_in = sum(value[component == "internal_in"]) + sum(value[component == "international_in"]),
            total_out = sum(value[component == "internal_out"]) + sum(value[component == "international_out"]),
            total_net = sum(value[component == "internal_net"]) + sum(value[component == "international_net"]),
            .groups = "drop") %>%
  pivot_longer(cols = c("total_in", "total_out", "total_net"), names_to = "component", values_to = "value")


dpm_comp_ratios <- dpm_comp %>%
  filter(component %in% c("international_in", "international_out",
                          "internal_in", "internal_out")) %>%
  pivot_wider(names_from = component, values_from = value) %>%
  mutate(inflow = internal_in/(international_in + internal_in),
         outflow = internal_out/(international_out + internal_out)) %>%
  select(-c(internal_in, international_in, internal_out, international_out)) %>%
  pivot_longer(cols = c("inflow", "outflow"), names_to = "direction", values_to = "proportion_internal")


dpm_comp_totals <- dpm_comp %>%
  filter(!component %in% c("total_in", "total_out", "total_net")) %>%
  bind_rows(dpm_total_flow_comps) %>%
  group_by(across(-any_of(c("value", "age")))) %>%
  summarise(value = sum(value), .groups = "drop") %>%
  arrange(year) %>%
  group_by(across(-any_of(c("value", "year")))) %>%
  mutate(cumulative_value = cumsum(value)) %>%
  ungroup()

```


# convert_upc

This repository contains two related examples of:

* how the Unattributable Population Change (UPC) component included in the rebased ONS mid-year estimate (MYE) population back series could be replaced with adjustments to the existing annual international migration estimates.

* how total in- and out-flows in ONS's current Admin-Based Population Estimates (ABPE)/Dynamic Population Model (DPM) estimates could be split into separate international and domestic flows.

## Converting UPC in the rebased MYE series

The method shown here uses a single function (*R/functions/optimise_gross_flows.R*) that takes an initial pair gross flows, together with a target net value, and returns new gross flows that are consistent with this net figure and represent the minimum 'cost' change to the original flows. 

This function was originally developed by the GLA in its process for creating the [rebased population estimates](https://data.london.gov.uk/dataset/modelled-population-backseries) used as a basis for the GLA's population projections.  

The full GLA rebasing process is somewhat complex and involves fitting new annual net flows before splitting these out into gross flows and building a consistent annual population series. 

However, if the goal is just to reassign already calculated UPC to international flows, without modifying the estimated population, then the process is greatly simplified.

## Splitting ABPE total flows

This method uses a related function (*R/functions/split_gross_flows.R*) to create modelled international and domestic gross flow  estimates, based on a pair of initial estimates of each, and consistent with a 'target' total flow figure, while accounting for the user's relative confidence in the accuracy of these base estimates.

Here the base domestic and international flows are taken from the ONS rebased MYE series and used to split the total flow components in the published ABPE/DPM series.

Confidence in the accuracy of flow estimates will vary by area, age, sex, direction, and year. The function allows the user to readily account for her understanding of the relative accuracy of the base estimates across all these dimensions, but the example shown here uses largely arbitrary values of the relative confidence parameters for the purpose of illustration. 

## Instructions

The two processes are run from the scripts *R/run_process_ons2023_mye_series.R* and *R/run_process_ons_dpm_series.R*, respectively.

*R/run_process_ons2023_mye_series.R* will:

* Fetch and clean the detailed mid-year estimates series published by ONS that covers the period 2011 to 2023 

* Create modelled alternative annual international gross flows that are consistent with the sum of the international_net and UPC components

* Write out the new series to *data/processed/* as both an RDS file in tidy format and as a csv in a similar format to that originally published by ONS

*R/run_process_ons_dpm_series.R* will:

* Fetch and clean the detailed mid-year estimates series published by ONS that covers the period 2011 to 2023 

* Fetch and clean the current admin-based estimates series published July 2024, which also covers the period 2011 to 2023 

* Create modelled annual international and internal flow components that are consistent with the original total immigration and emigration components

* Write out the new series to *data/processed/* as both an RDS file in tidy format and as a csv in a similar format to that originally published by ONS

### Required packages

dplyr, tidyr, readxl, readr, stringr


# Overview of methodologies

## Converting UPC in the rebased MYE series

In population modelling, it is often the case that we wish to make an adjustment to the size of a population stock or net migration flow and must make corresponding adjustment to the size of the underlying gross flows, such that they remain consistent. When the adjustment to the net flow is positive, consistent gross flows can be created by some combination of higher inflows and/or lower outflows (and vice versa). As a given net flow can be described by any number of different pairs of in- and out-flows, the problem is to choose the most appropriate values for the in- and outflows.

The approach taken by the GLA to this problem is to try and identify adjustments that minimise (from a statistical perspective) changes from the original gross flow estimates.  

* Gross flows are treated as having an underlying probability density described by a Poission distribution with means equal to the original estimates. 

* Starting from the initial values of the gross flows, an incremental adjustment is made to either the inflow or outflow according to which would see the smaller decrease in the probability density after the adjustment is made.  

* Successive incremental adjustments are made until the gross flows are consistent with desired net flow.

The relative size of the adjustments to the in- and out-flows is determined by the sizes of the initial flows and the size and direction of the adjustment required.  In cases where the initial gross flows are already consistent with the target net, the function will return these values unchanged. 

The example below illustrates how the modelled gross flows for a given pair of base flows vary with the target net figure.

```{r, echo = FALSE}

example_df <- data.frame(target_net = c(-100:100), base_in = 40, base_out = 15) %>%
  mutate(model_flows = optimise_gross_flows(base_in, base_out, target_net)) %>%
  unnest_wider(col = model_flows) %>%
  mutate(outflow = -outflow) %>%
  pivot_longer(cols = c("inflow", "outflow"), names_to = "direction", values_to = "gross_flows")

example_df %>%
  ggplot(aes(x = target_net, y = gross_flows, colour = direction)) +
  theme_minimal() +
  annotate("point", x = c(25, 25), y = c(40, -15)) +
  geom_hline(yintercept = 0, colour = "darkgrey") +
  geom_line() +
  labs(title = "Returned gross flow values by net flow target",
       subtitle = "base inflow = 40, base outflow = 15, original net = +25")


```

```{r, include=FALSE}

sel_yr <- 2019
sel_nm <- "Camden"
sel_sex <- "female"
age_max <- 50

```


In this application the target international net flow is taken to be the sum of the values of the original international net flow and the UPC component.

The following charts illustrate the original and adjusted data for females in Camden - a group with a very high UPC component in the rebased estimates which removes almost 18 thousand persons from the population over the course of the decade.   


```{r, echo = FALSE}

mye_comp_totals %>%
      filter(gss_name == sel_nm,
         sex == sel_sex) %>%
  filter(component %in% c("international_net", "unattrib")) %>%
  filter(between(year, 2011, 2021)) %>%
  ggplot(aes(x = year, y = value, colour = source, linetype = component)) +
  theme_minimal() +
  geom_hline(yintercept = 0, colour = "darkgrey") +
  geom_line() +
  labs(title = "Original and adjusted annual international net and UPC",
       subtitle = paste0(sel_nm, ", ", sel_sex)) +
    scale_x_continuous(n.breaks = 9)

```

```{r, echo = FALSE}

mye_comp_totals %>%
    filter(gss_name == sel_nm,
         sex == sel_sex) %>%
  filter(component %in% c("international_net", "unattrib")) %>%
  filter(between(year, 2011, 2021)) %>%
  ggplot(aes(x = year, y = cumulative_value, colour = source, linetype = component)) +
  theme_minimal() +
  geom_hline(yintercept = 0, colour = "darkgrey") +
  geom_line() +
  labs(title = "Cummulative contribution of international net and UPC",
       subtitle = paste0(sel_nm, ", ", sel_sex)) +
    scale_x_continuous(n.breaks = 9)

```

```{r, echo = FALSE}

mye_comp %>%
  filter(gss_name == sel_nm,
         sex == sel_sex) %>%
  filter(year == sel_yr) %>%
   filter(age <= age_max) %>%
  filter(component %in% c("international_net", "unattrib")) %>%
  ggplot(aes(x = age, y = value, colour = source, linetype = component)) +
  theme_minimal() +
  geom_hline(yintercept = 0, colour = "darkgrey") +
  geom_line() +
  labs(title = "Original and adjusted international net and UPC",
  subtitle = paste0(sel_nm, ", ", sel_sex, ", ", sel_yr, ", age 0 to ", age_max)) +  
  scale_x_continuous(n.breaks = 9) +
  facet_wrap("source", ncol = 1)

```

As intended, the adjusted gross flows closely track the distributions of the original estimates by year and by age.

```{r, echo = FALSE}

mye_comp_totals %>%
  filter(gss_name == sel_nm,
         sex == sel_sex) %>%
  filter(component %in% c("international_in", "international_out", "international_net")) %>%
  filter(between(year, 2011, 2021)) %>%
  ggplot(aes(x = year, y = value, colour = component, linetype = source)) +
  theme_minimal() +
  geom_hline(yintercept = 0, colour = "darkgrey") +
  geom_line() +
  labs(title = "Original and adjusted annual international gross flows",
       subtitle = paste0(sel_nm, ", ", sel_sex)) +
    scale_x_continuous(n.breaks = 9)

```

```{r, echo = FALSE}

mye_comp %>%
  filter(gss_name == sel_nm,
         sex == sel_sex) %>%
  filter(year == sel_yr) %>%
   filter(age <= age_max) %>%
  filter(component %in% c("international_in", "international_out")) %>%
  ggplot(aes(x = age, y = value, colour = component, linetype = source)) +
  theme_minimal() +
  geom_hline(yintercept = 0, colour = "darkgrey") +
  geom_line() +
  labs(title = "Original and adjusted international gross flows by age",
         subtitle = paste0(sel_nm, ", ", sel_sex, ", ", sel_yr, ", age 0 to ", age_max)) +  
    scale_x_continuous(n.breaks = 9) +
  facet_wrap("component", ncol = 1)

```

## Splitting total gross flows from the DPM outputs

ONS has stated that it has a goal of making its admin-based population estimates the official series from 2025. To date, experimental outputs from the Dynamic Population Model - used to produce the admin-based estimates - do not include breakdowns of gross flows into separate domestic and international components.  While outputs are eventually expected to include the full set of components included in the current official series, the GLA has an interest in understanding how switching to the admin-based estimates may impact on the results of population projections so that the upcoming change can be can be accounted for in its work programme. Here we consider one option for splitting the total immigration and total emigration flows into international and domestic components, so that the data can be used as inputs to existing population models.  

The approach makes use of the domestic and international flows from the rebased MYE series to provide a basis for dividing the ABPE total flows. The simplest way to split the flows would likely be to simply allocate the total flows between international and domestic in the same proportions as in the MYE. The function used here supports this approach, but also allows the split to account for the relative confidence the user has in the accuracy of the original international and domestic flow estimates. If, for example, the user specifies that they have lower confidence in the international than the domestic estimates, the function will account for this in the fitting process - allowing the modelled international flow to more readily deviate from the base MYE value than it does the domestic when matching the total flow.

The plot below illustrates how, for a given pair of base flows, the modelled international and domestic flows vary with both the target total flow figure and the relative confidence parameter specified, where:

* a value of 1 represents equal confidence in the original international and domestic flow estimates and results in the same proportional split as the base flows
* values < 1 represent lower levels of confidence in the accuracy of the original international flow estimates
* values > 1 represent lower levels of confidence in the accuracy of the original domestic flow estimates


```{r, echo = FALSE}

example_df <- expand.grid(target_total = c(0:400), relative_international_confidence = c(0.2, 0.5, 0.9, 1.1, 2, 5), 
                         base_international = 50, base_domestic = 100) %>%
  mutate(model_flows = split_gross_flows(base_international, base_domestic, target_total,
                                         relative_international_confidence,
                                         jump_scale = 30)) %>%
  unnest_wider(col = model_flows) %>%
  select(-c(base_international, base_domestic)) %>%
  pivot_longer(cols = c("new_international", "new_domestic"), names_to = "flow", values_to = "value")

example_df %>%
  ggplot(aes(x = target_total, y = value, colour = flow)) +
  theme_minimal() +
  geom_hline(yintercept = 0, colour = "grey", alpha = 0.5) +
  annotate("point", x = c(150, 150), y = c(50, 100)) +
  geom_line() +
  labs(title = "Returned flows by total target and choice of relative confidence parameter",
       subtitle = "base international = 50, base domestic = 100, original total flow = 150") +
  facet_wrap("relative_international_confidence", ncol = 3)


```

```{r, include=FALSE}

sel_yr <- 2022
sel_nm <- "Camden"
sel_sex <- "female"
age_max <- 50

```

The following charts illustrate the modelled flows for the DPM alongside the original flows from the official mid-year estimates for females in Camden.  In this example, the relative confidence parameter for inflows has been (arbitrarily) set at 3 for inflows and 0.3 for outflows, meaning that, effectively, domestic migration will be prioritised for adjustment when fitting to total inflows and international migration will be prioritised for adjustment when fitting outflows.  In this case, the relative confidence parameters do not vary by age/sex, but there is scope to do so to account for known variations in the quality of the data for particular population groups.


```{r, echo = FALSE}

dpm_comp_totals %>%
  filter(gss_name == sel_nm,
         sex == sel_sex) %>%
  filter(component %in% c("total_in", "total_out", "total_net")) %>%
  filter(source %in% c("DPM", "MYE original")) %>%
  filter(between(year, 2011, 2023)) %>%
  ggplot(aes(x = year, y = value, colour = component, linetype = source)) +
  theme_minimal() +
  geom_hline(yintercept = 0, colour = "darkgrey") +
  geom_line() +
  labs(title = "Comparison of MYE and DPM annual total flows",
       subtitle = paste0(sel_nm, ", ", sel_sex)) +
    scale_x_continuous(n.breaks = 9)

```


```{r, echo = FALSE}

dpm_comp_totals %>%
  filter(gss_name == sel_nm,
         sex == sel_sex) %>%
  filter(component %in% c("international_net", "internal_net", "total_net")) %>%
  filter(source %in% c("DPM", "MYE original")) %>%
  filter(between(year, 2011, 2023)) %>%
  ggplot(aes(x = year, y = value, colour = component, linetype = source)) +
  theme_minimal() +
  geom_hline(yintercept = 0, colour = "darkgrey") +
  geom_line() +
  labs(title = "Comparison of MYE and modelled DPM net flows",
       subtitle = paste0(sel_nm, ", ", sel_sex)) +
    scale_x_continuous(n.breaks = 9)

```

```{r, echo = FALSE}

dpm_comp_totals %>%
  filter(gss_name == sel_nm,
         sex == sel_sex) %>%
  filter(component %in% c("international_in", "international_out", "international_net")) %>%
  filter(source %in% c("DPM", "MYE original")) %>%
  filter(between(year, 2011, 2023)) %>%
  ggplot(aes(x = year, y = value, colour = component, linetype = source)) +
  theme_minimal() +
  geom_hline(yintercept = 0, colour = "darkgrey") +
  geom_line() +
  labs(title = "Comparison of MYE and modelled DPM international flows",
       subtitle = paste0(sel_nm, ", ", sel_sex)) +
    scale_x_continuous(n.breaks = 9)

```

```{r, echo = FALSE}

dpm_comp_totals %>%
  filter(gss_name == sel_nm,
         sex == sel_sex) %>%
  filter(component %in% c("internal_in", "internal_out", "internal_net")) %>%
  filter(source %in% c("DPM", "MYE original")) %>%
  filter(between(year, 2011, 2023)) %>%
  ggplot(aes(x = year, y = value, colour = component, linetype = source)) +
  theme_minimal() +
  geom_hline(yintercept = 0, colour = "darkgrey") +
  geom_line() +
  labs(title = "Comparison of MYE and modelled DPM internal flows",
       subtitle = paste0(sel_nm, ", ", sel_sex)) +
    scale_x_continuous(n.breaks = 9)

```

```{r, echo = FALSE}

dpm_comp %>%
  bind_rows(dpm_total_flow_comps) %>%
  filter(gss_name == sel_nm,
         sex == sel_sex) %>%
  filter(year == sel_yr) %>%
   filter(age <= age_max) %>%
  filter(component %in% c("international_net", "internal_net", "total_net")) %>%
  filter(source %in% c("DPM", "MYE original")) %>%
  ggplot(aes(x = age, y = value, colour = component, linetype = source)) +
  theme_minimal() +
  geom_hline(yintercept = 0, colour = "darkgrey") +
  geom_line() +
  labs(title = "Comparison of MYE and modelled DPM net flows by age",
       subtitle = paste0(sel_nm, ", ", sel_sex, ", ", sel_yr, ", age 0 to ", age_max)) +
  scale_x_continuous(n.breaks = 9) +
  facet_wrap("component", scales = "free_y", ncol = 1)

```

```{r, echo = FALSE}

dpm_comp %>%
  bind_rows(dpm_total_flow_comps) %>%
  filter(gss_name == sel_nm,
         sex == sel_sex) %>%
  filter(year == sel_yr) %>%
   filter(age <= age_max) %>%
  filter(component %in% c("total_in", "total_out")) %>%
  filter(source %in% c("DPM", "MYE original")) %>%
  ggplot(aes(x = age, y = value, colour = component, linetype = source)) +
  theme_minimal() +
  geom_hline(yintercept = 0, colour = "darkgrey") +
  geom_line() +
  labs(title = "Comparison of MYE and DPM total gross flows by age",
       subtitle = paste0(sel_nm, ", ", sel_sex, ", ", sel_yr, ", age 0 to ", age_max)) +
    scale_x_continuous(n.breaks = 9) +
    facet_wrap("component", ncol = 1)


```


```{r, echo = FALSE}

dpm_comp %>%
  filter(gss_name == sel_nm,
         sex == sel_sex) %>%
  filter(year == sel_yr) %>%
  filter(age <= age_max) %>%
  filter(component %in% c("internal_in", "internal_out", "international_in", "international_out")) %>%
  filter(source %in% c("DPM", "MYE original")) %>%
  ggplot(aes(x = age, y = value, colour = component, linetype = source)) +
  theme_minimal() +
  geom_hline(yintercept = 0, colour = "darkgrey") +
  geom_line() +
  labs(title = "",
       subtitle = paste0(sel_nm, ", ", sel_sex, ", ", sel_yr, ", age 0 to ", age_max)) +
  scale_x_continuous(n.breaks = 9) +
  facet_wrap("component", ncol = 2, scales = "free_y")

```
